### 2023-07-03 chapters/jet_substructure.tex
Adjusted chapters, new jet substructure chapter.

\chapter{Jet Substructure}

Quantifying the internal structure of hadronic jets plays an important role in translating between experimental results and theoretical predictions.
%
The toolkit of jet substructure, including techniques from jet grooming \cite{Krohn:2009th,Ellis:2009me,Larkoski:2014wba,Dasgupta:2013ihk,Dasgupta:2013via,Tseng:2013dva,Dasgupta:2016ktv} and pile-up mitigation \cite{Cacciari:2007fd,Soyez:2012hv,Komiske:2017ubm,Bertolini:2014bba,Cacciari:2008gn,Butterworth:2008iy,Chatrchyan:2012sn,Alon:2011xb,Krohn:2013lba,ATLAS-CONF-2013-083,ATLAS-CONF-2013-085,Chatrchyan:2012tt,CMS-PAS-JME-13-005,Sirunyan:2020foa,ATLAS-CONF-2012-066,Monk:2018clo,Martinez:2018fwc},
helps characterize the radiation patterns produced by jets in a wide variety of physical processes at colliders.
%
Jet substructure can distinguish jets produced by light quarks, gluons, and the decays of electroweak bosons and top quarks
%
\cite{Thaler:2008ju,Thaler:2011gf,Hook:2011cq,Gallicchio:2011xq,Soper:2012pb,Gallicchio:2012ez,CMS-PAS-JME-09-001,CMS-PAS-EXO-09-002,CMS:2013kfa,ATL-PHYS-PUB-2009-081,ATL-PHYS-PUB-2010-008,Cui:2010km,ATLAS-CONF-2011-053,Chatrchyan:2013rla,Larkoski:2013eya,Dasgupta:2012hg,Backovic:2013bga,ATLAS-CONF-2013-084,Komiske:2018vkc,Komiske:2016rsd,Metodiev:2018ftz,Krohn:2012fg,MERINO:2013tta,Bhattacherjee:2016bpy,Macaluso:2018tck,Egan:2017ojy,Kasieczka:2017nvn,Pearkes:2017hku,Butter:2017cot},
%
provide accurate predictions of the behavior of boosted objects in many contexts
%
\cite{Catani:1992ua,Dokshitzer:1998kz,Dasgupta:2001sh,Banfi:2004yd,Banfi:2005gj,Ellis:2009su,Banfi:2010pa,Walsh:2011fz,Chien:2012ur,Li:2012bw,Jouttenus:2013hs,Hatta:2013iba,Larkoski:2014tva,Procura:2018zpn,Aaboud:2017aca,Frye:2016aiz, Almeida:2008yp,Larkoski:2017iuy,Larkoski:2017cqq,Thaler:2010tr,Ellis:2009me,Abdesselam:2010pt,Katz:2010mr,Gallicchio:2010dq,Adams:2015hiv,Sirunyan:2017ezt,Moore:2018lsr,FerreiradeLima:2016gcz,Rubin:2010fc,Chatrchyan:2012sn,ATLAS:2019kwg,CMS-PAS-BTV-13-001,CMS-PAS-JME-13-006,Kribs:2009yh,Chen:2010wk,Hackstein:2010wk,Kim:2010uj,Almeida:2011aa,Pandolfi:2012ima,Vernieri:2014wfa,CMS-PAS-HIG-17-007,Procura:2014cba},
%
and even reveal non-perturbative features of the infrared structure of QCD
%
\cite{Komiske:2019jim, Komiske:2019fks, Hoang:2019ceu, Mateu:2012nk, Wobisch:1998wt}.

%--------------------------------
% Soft Distortion and Additive Contamination
%--------------------------------
Jet grooming facilitates our understanding of jet substructure by removing low-energy pollution from jets.
%
The major goal of jet grooming is to reduce the sensitivity of jet observables to two conceptually distinct categories of pollution:
\begin{itemize}
    \item
    \textbf{Soft Distortions}
    are low-energy effects that distort the radiation coming directly from hard jet production, such as hadronization and detector effects;

    \item
    \textbf{Additive Contamination}
    denotes a backdrop of low-energy radiation that pollutes a jet but is not directly connected to jet production, including the pileup (PU) of radiation from overlapping particle collisions and an underlying event (UE) of secondary parton interactions.
\end{itemize}
The low-energy phenomena of QCD which produce both soft contamination and additive contamination are notoriously difficult to model, and jet grooming has become an important toolkit used to empower both experimental and theoretical analyses of jet substructure
\cite{Krohn:2009th,Ellis:2009me,Larkoski:2014wba,Dasgupta:2013ihk,Dasgupta:2013via,Tseng:2013dva,Dasgupta:2016ktv,Thaler:2008ju,Thaler:2011gf,Hook:2011cq,Gallicchio:2011xq,Soper:2012pb,Gallicchio:2012ez,CMS-PAS-JME-09-001,CMS-PAS-EXO-09-002,CMS:2013kfa,ATL-PHYS-PUB-2009-081,ATL-PHYS-PUB-2010-008,Cui:2010km,ATLAS-CONF-2011-053,Chatrchyan:2013rla,Larkoski:2013eya,Dasgupta:2012hg,Backovic:2013bga,ATLAS-CONF-2013-084,Komiske:2018vkc,Komiske:2016rsd,Metodiev:2018ftz,Krohn:2012fg,MERINO:2013tta,Bhattacherjee:2016bpy,Macaluso:2018tck,Egan:2017ojy,Kasieczka:2017nvn,Pearkes:2017hku,Butter:2017cot,Catani:1992ua,Dokshitzer:1998kz,Dasgupta:2001sh,Banfi:2004yd,Banfi:2005gj,Ellis:2009su,Banfi:2010pa,Walsh:2011fz,Chien:2012ur,Li:2012bw,Jouttenus:2013hs,Hatta:2013iba,Larkoski:2014tva,Procura:2018zpn,Aaboud:2017aca,Frye:2016aiz, Almeida:2008yp,Larkoski:2017iuy,Larkoski:2017cqq,Thaler:2010tr,Abdesselam:2010pt,Katz:2010mr,Gallicchio:2010dq,Adams:2015hiv,Sirunyan:2017ezt,Moore:2018lsr,FerreiradeLima:2016gcz,Rubin:2010fc,Chatrchyan:2012sn,ATLAS:2019kwg,CMS-PAS-BTV-13-001,CMS-PAS-JME-13-006,Kribs:2009yh,Chen:2010wk,Hackstein:2010wk,Kim:2010uj,Almeida:2011aa,Pandolfi:2012ima,Vernieri:2014wfa,CMS-PAS-HIG-17-007,Procura:2014cba,ATL-PHYS-PUB-2019-027,Aad:2019vyi,ATLAS:2020gwe}.

A complication of many popular jet grooming procedures is that they are \textit{discontinuous} in event space:
%
they may map similar ungroomed jets, which differ only by small changes in their distributions of energy, into extensively distinct groomed jets \cite{Dasgupta:2013ihk,Larkoski:2014wba}.
%
Discontinuous behavior in jet grooming algorithms leads in turn to undesirably sensitive responses to low-energy pollution.
%
Experimentally, discontinuities in grooming lead to large responses of groomed substructure to the interactions of jets with experimental detectors, and resulting difficulties in unfolding \cite{ATL-PHYS-PUB-2019-027,Aad:2019vyi,ATLAS:2020gwe}.
%
Theoretically, discontinuities in grooming lead to uncertainties in fixed-order calculations \cite{Larkoski:2014wba} and complications in describing the effects of hadronization on groomed jet substructure \cite{Hoang:2019ceu}.

%=======================
% Introducing \PIRANHA{}: In this paper we...
%=======================
In this paper, we introduce \textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{n}i\textbf{h}il\textbf{a}tion (\PIRANHA{}), a paradigm for continuous jet grooming, using recent geometric perspectives in collider physics.
%
\PIRANHA{} applies intuition from optimal transport theory and the Energy Mover's Distance (EMD) between collider events~\cite{Komiske:2019fks,Komiske:2020qhg} to overcome the discontinuous behavior of existing jet grooming algorithms.
%
We introduce the new terminology of \textit{soft} and \textit{angular} \textit{discontinuities} to categorize discontinuous behavior under infinitesimal changes to the energies and angles of jet constituents, respectively.
%
We explore the improved continuity of \PIRANHA{} conceptually through several toy examples, and more practically by studying the responses of \PIRANHA{} grooming to several sources of soft distortions and additive contamination.

At present, \PIRANHA{} has three concrete implementations:
\begin{itemize}
    \item
    \textbf{Apollonius Subtraction} (P-AS), introduced in \Reff{Komiske:2020qhg}, is a continuous, conceptually simple application of the \PIRANHA{} paradigm with a computationally expensive implementation;

    \item
    \textbf{Iterated Voronoi Subtraction} (P-IVS), also introduced in \Reff{Komiske:2020qhg}, is a continuous, computationally efficient modification of P-AS with a simple geometric implementation;

    \item
    \textbf{Recursive Subtraction} (P-RS) is a family of nearly-continuous, computationally efficient, tree-based algorithms motivated by P-AS and P-IVS.
    %
    We develop the concrete implementation of \textbf{Recursive Subtraction with a Fraction \(\boldsymbol{f}\)} (P-RSF\(_f\)).
    %
    We discuss how P-RSF\(_f\) is discontinuous in suppressed regions of parameter space, and focus on \PRSF{1/2}, which is the only fully soft-continuous version of P-RSF\(_f\).
\end{itemize}

\Tab{groomerlist} expresses the infrared/collinear safety (using the definition of \Reff{Komiske:2020qhg}) and soft/angular continuity properties of the traditional and \PIRANHA{} groomers we discuss in this paper.
%
The soft continuity properties expressed in \Tab{groomerlist} are simple to understand by comparing the hard-cutoff framework of popular groomers to the subtractive framework of \PIRANHA{}.
%
The angular discontinuities expressed in \Tab{groomerlist} are more subtle and are inherited from angular-ordered jet clustering.
%
We note that the Recursive Subtraction algorithms we introduce are discontinuous only in a highly suppressed region of phase space, and we still find it suitable to call P-RS a family of \PIRANHA{} groomers.

\begin{table}[t]
\centering
\begin{tabular}{|c|c||c|c||c|c|}
\hhline{~~--||--}
\multicolumn{2}{c|}{}
&
\multicolumn{2}{c||}{Safety}
&
\multicolumn{2}{c|}{Continuity}
\\
\hhline{~~--||--}
\noalign{\vskip\doublerulesep
         \vskip-\arrayrulewidth}
\hhline{--||--||--}
    Groomer & Acronym &
    Infrared & Collinear &
    Soft & Angular
\\
\hhline{--||--||--}
\noalign{\vskip\doublerulesep
         \vskip-\arrayrulewidth}
\hhline{--||--||--}
Apollonius Subtraction & P-AS &
\cmark & \cmark &
\cmark & \cmark
\\
\hhline{--||--||--}
Iterated Voronoi Subtraction & P-IVS &
\cmark & \cmark &
\cmark & \cmark
\\
\hhline{--||--||--}
Recursive Sub. w/Fraction \(f = 1/2\) & \PRSF{1/2} &
\cmark & \cmark &
\cmark & \danger
\\
\hhline{--||--||--}
Recursive Sub. w/Fraction \(f \neq 1/2\) &  P-RSF\(_{f \neq 1/2}\) &
\cmark & \cmark &
\danger & \danger
\\
\hhline{--||--||--}
\noalign{\vskip\doublerulesep
         \vskip-\arrayrulewidth}
\hhline{--||--||--}
Soft Drop, \(\beta_{\rm SD} > 0\) & SD\(_{\beta > 0}\) &
\cmark & \cmark &
\xmark & \danger
\\
\hhline{--||--||--}
Soft Drop, \(\beta_{\rm SD} \leq 0\) & SD\(_{\beta \leq 0}\) &
\cmark & \xmark &
\xmark & \danger
\\
\hline
\end{tabular}
\caption{
Safety and continuity properties of the \PIRANHA{} and Soft Drop grooming algorithms studied in this paper.
%
The mark \cmark~indicates that a groomer satisfies the associated condition, while the mark \xmark~indicates that it does not.
%
The mark \raisebox{0.15 em}{\danger} indicates that a groomer does not satisfy the condition, but only in a boundary region of phase space which is exponentially suppressed at leading order in perturbative QCD.
}
\label{tab:groomerlist}
\end{table}

%-----------------------------------
% Discontinuity visualization.:
%-----------------------------------
\begin{figure}[tp]
\centering
\scalebox{0.95}{
\begin{tikzpicture}[baseline=-3.5ex]

% Event Labels
  \node at (-1.7,  .6)()
        {\huge \(\mathcal{E}^+\)};
  \node at (-1.7,  -1.25)()
        {\huge\(\mathcal{E}^-\)};

% Header wrapper
\draw [fill=black, opacity=0.05]
       (-0.7,2.4) -- (13.6,2.4) -- (13.6,1.65) -- (-0.7,1.65) -- cycle;

%%%%%%%%%%%%%%
% Ungroomed
%%%%%%%%%%%%%%
  \node at (1.5,  2)() {\textbf{Ungroomed}};

% E+
  % points
  \coordinate (A)  at (1.4, 1.4);
  \coordinate (O)  at (0, 0);
  \coordinate (B)  at (3, 0);
  % legs
  \draw[-Stealth,line width=.55mm] (O) -> (A);
  \draw[-Stealth,line width=.7mm] (O) -> (B);

% E-
  % points
  \coordinate (A)  at (.8, -1.2);
  \coordinate (O)  at (0, -2);
  \coordinate (B)  at (3, -2);
  % legs
  \draw[-Stealth,line width=.55mm] (O) -> (A);
  \draw[-Stealth,line width=.7mm] (O) -> (B);

% lines

% zcut lines
\coordinate (I)  at (0.7, 1.0);
\coordinate (F)  at (3.15, 1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);
\node at (3.5,  1.0)()
        {\(\zcut\)};

\coordinate (I)  at (0.7, -1.0);
\coordinate (F)  at (3.15, -1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);
\node at (3.5,  -1.0)()
        {\(\zcut\)};

% Descriptors
\node [rotate=45] at (0.0,  0.5)()
        {\(z^+ > \zcut\)};
\node [rotate=45] at (0.0,  -1.5)()
        {\(z^- < \zcut\)};

%%%%%%%%%%%%%%
% Soft Drop Grooming
%%%%%%%%%%%%%%
\node at (6.5,  2)() {\textcolor{dodgerblue!75!black}{\textbf{Soft Drop}, \(\beta_{\rm SD} = 0\)}};

% E+
  % points
  \coordinate (A)  at (6.4, 1.4);
  \coordinate (O)  at (5.0, 0);
  \coordinate (B)  at (8.0, 0);
  % legs
  \draw[-Stealth,line width=.55mm, color=royalblue] (O) -> (A);
  \draw[-Stealth,line width=.7mm, color=royalblue] (O) -> (B);

% E-
  % points
  \coordinate (Ap)  at (5.8, -1.2);
  \coordinate (O)  at (5.0, -2);
  \coordinate (B)  at (8.0, -2);
  % legs
  \draw[-Stealth, line width=.55mm, color=black!25!white] (O) -> (Ap);
  \draw[-Stealth,line width=.7mm, color=royalblue] (O) -> (B);

% lines
% zcut lines
\coordinate (I)  at (5.7, 1.0);
\coordinate (F)  at (8.15, 1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);

\coordinate (I)  at (5.7, -1.0);
\coordinate (F)  at (8.15, -1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);

%%%%%%%%%%%%%%
% PIRANHA Grooming \PRSF{1/2}
%%%%%%%%%%%%%%
  \node at (10.9,  2)() {\textcolor{ochre!75!black}{\textbf{P-RSF} (arb.\,\(f_{\rm soft}\))}};

% E+
  % points
  \coordinate (A)  at (10.4, 0.9);
  \coordinate (Ap)  at (10.9, 1.4);
  \coordinate (O)  at (9.5, 0);
  \coordinate (B)  at (11.7, 0);
  \coordinate (Bp)  at (12.5, 0);
  % legs
  \draw[-Stealth,line width=.55mm, color=black!25!white] (O) -> (Ap);
  \draw[-Stealth,line width=.55mm, color=darkgoldenrod] (O) -> (A);
  \draw[-Stealth,line width=.7mm, color=black!25!white] (O) -> (Bp);
  \draw[-Stealth,line width=.7mm, color=darkgoldenrod] (O) -> (B);

% E-
  % points
  \coordinate (A)  at (9.9, -1.6);
  \coordinate (Ap)  at (10.3, -1.2);
  \coordinate (O)  at (9.5, -2);
  \coordinate (B)  at (11.7, -2);
  \coordinate (Bp)  at (12.5, -2);
  % legs
  \draw[-Stealth,line width=.55mm, color=black!25!white] (O) -> (Ap);
  \draw[-Stealth,line width=.55mm, color=darkgoldenrod] (O) -> (A);
  \draw[-Stealth,line width=.7mm, color=black!25!white] (O) -> (Bp);
  \draw[-Stealth,line width=.7mm, color=darkgoldenrod] (O) -> (B);

% lines
% zcut lines
\coordinate (I)  at (10.2, 1.0);
\coordinate (F)  at (12.65, 1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);

\coordinate (I)  at (10.2, -1.0);
\coordinate (F)  at (12.65, -1.0);
\draw[dash pattern=on 2pt off 3pt, line width=.4mm, color=aurometalsaurus] (I) -> (F);

%%%%%%%%%%%%%%
% QCD
%%%%%%%%%%%%%%
% Double line
\coordinate (I)  at (-2.5, -2.8);
\coordinate (F)  at (13.3, -2.8);
\draw[line width=.3mm] (I) -> (F);

\coordinate (I)  at (-2.5, -2.9);
\coordinate (F)  at (13.3, -2.9);
\draw[line width=.3mm] (I) -> (F);

% Labels
  \node at (-1.7,  -4.6)()
        {\textbf{\LARGE QCD}};
  \node at (-2.55, -5.8)(){\Large (};
  \node at (-1.7,  -5.6)()
        {\texttt{Pythia}};
  \node at (-1.7,  -6.0)()
        {\texttt{8.244}};
 \node at (-0.85, -5.8)(){\Large )};
  \node at (-1.7,  -8.4)()
         {\textbf{\LARGE QCD}};
 \node at (-1.7,  -9.0)()
         {\textbf{\Large +}};
 \node at (-1.7,  -9.6)()
        {\textbf{\LARGE Noise}};

% Event visualization
% Ungroomed
  \node at (1.8,  -5.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_raw_1}};
  \node at (1.8,  -9.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_raw_2}};
% mMDT
  \node at (6.5,  -5.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_mmdt_1}};
  \node at (6.5,  -9.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_mmdt_2}};
% P-RSF_1/2
  \node at (11.3,  -5.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_rsf_half_1}};
  \node at (11.3,  -9.0)() {\includegraphics[width=.25\textwidth]{figures/event_visualizations/vis_rsf_half_2}};

% Axis info
 \node at (-0.4,  -5.0)() {\Large \(\phi\)};
 \node at (-0.4,  -9.0)() {\Large \(\phi\)};
 \node at (1.8,  -11.2)() {\Large \(y\)};
 \node at (6.5,  -11.2)() {\Large \(y\)};
 \node at (11.3,  -11.2)() {\Large \(y\)};

 % phi labels
 \node at (-0.25,  -10.77)() {\footnotesize -1};
 \node at (-0.25,  -7.3)() {\footnotesize 1};
 \node at (-0.25,  -6.77)() {\footnotesize -1};
 \node at (-0.25,  -3.3)() {\footnotesize 1};
 % y labels
 \node at (-0.05,  -11.1)() {\footnotesize -1};
 \node at (3.65,  -11.1)() {\footnotesize 1};
 \node at (4.65,  -11.1)() {\footnotesize -1};
 \node at (8.35,  -11.1)() {\footnotesize 1};
 \node at (9.45,  -11.1)() {\footnotesize -1};
 \node at (13.15,  -11.1)() {\footnotesize 1};

 %\node at (1.8,  -5.0)() {-1.0};
\end{tikzpicture}
}
\caption{
A visualization of how similar ungroomed events (first column, in black) may be mapped discontinuously to distinct groomed jets by the hard-cutoff mMDT algorithm (Soft Drop with \(\beta_{\rm SD} = 0\), or \SD{0}; middle column, in blue), while the continuous \PRSF{1/2} algorithm will map them to similar groomed results (final column, in orange).
%
In the first two rows, we compare the action of mMDT and \PRSF{1/2} on simple toy events \(\mathcal{E}^\pm\), with particles visualized as arrows whose size corresponds to their energy.
%
In the final two rows, we compare the groomers acting on a more realistic QCD event produced with \texttt{Pythia 8.244}, and on the same QCD event with additional Gaussian noise, respectively.
}
\label{fig:intro_visual}
\end{figure}

%=======================
% Elaboration
%=======================
%--------------------------------
% Hard Cutoffs
%--------------------------------
Popular jet grooming algorithms are soft discontinuous because they introduce hard cutoffs on the energy of jet constituents, removing particles that are too low-energy to be classified as important probes of hard physics;
%
we call such groomers \textit{hard-cutoff groomers}.
%
The modified Mass Drop Tagger (mMDT) \cite{Dasgupta:2013ihk}, for example, requires that certain sub-jets within a jet carry a sufficient energy fraction, \(z > z_{\rm cut}\), to survive the grooming procedure.
%
If a sub-jet has an energy fraction near \(\zcut\), low-energy pollution may push its energy to the other side of the cutoff, leading to soft discontinuous behavior and associated theoretical and experimental complications.
%
\Fig{intro_visual} demonstrates this soft discontinuous behavior, the corresponding soft continuity of \PRSF{1/2}, by comparing their action on two sets of similar events:
%
two toy jet events \(\mathcal{E}^\pm\) consisting of two particles each, and a QCD jet produced in \texttt{Pythia 8.244} along with the same QCD event with a small amount of gaussian noise applied to the \(p_T\), rapidity, and azimuthal angle of each particle.
%
We revisit the events \(\mathcal{E}^\pm\) as tools to characterize soft discontinuity more precisely in \Sec{sd_discont}.
%
The mMDT grooming procedure maps \(\mathcal{E}^+\) and \(\mathcal{E}^-\) to very distinct events, and similarly maps the noiseless and noisy QCD jets to very distinct groomed jets.
%
We chose an unusually high value of \(\zcut = 0.25\) when grooming the QCD events to make the discontinuity of mMDT more apparent.
%
We will also use Soft Drop \cite{Larkoski:2014wba}, a generalization of mMDT reviewed in \Sec{softdrop} which reproduces mMDT when a parameter \(\beta_{\rm SD}\) is set to zero, as another example of soft discontinuous hard-cutoff grooming.

%--------------------------------
% \PIRANHA{}
%--------------------------------
\PIRANHA{} borrows technology from optimal transport to produce soft continuous, subtractive jet grooming algorithms without hard cutoffs.
%
\PIRANHA{} groomers may be roughly imagined as optimally transporting hungry piranhas to eat up delicious low-energy pollution.
%
Given an ungroomed event with energy density \(\mathcal{E}_{0}\) in the pseudorapidity-azimuth angle plane and any ansatz \(\rhoC\) for the energy density of low-energy pollution, one particularly succinct \PIRANHA{} algorithm is
\begin{align}
    \mathcal{E}_g\left[\mathcal{E}_0\right]
    =
    \underset{\mathcal{E}'}{\arg \min}\ {\rm EMD}_{\beta, R}
    \left(\mathcal{E}_0, \mathcal{E}' + \rhoC\right)
    \label{eqn:simple_piranha_intro}
    ,
\end{align}
where \(\mathcal{E}_g\) is a groomed energy distribution, and explicit dependence on the EMD metric makes the continuity of \(\mathcal{E}_g\left[\mathcal{E}_0\right]\) manifest.
%
P-AS is a direct, continuous application of \Eq{simple_piranha_intro} with \(\beta = 1\), and for which \(\rhoC\) is uniform in the pseudorapidity-azimuth plane to approximate common sources of additive contamination such as PU and UE \cite{Soyez:2018opl,Monk:2018clo,Sjostrand:1987su,Sjostrand:2014zea,Dasgupta:2007wa,Kirchgaesser:2020poq,Moraes:2007rq,CDF:2015txs,Larkoski:2021hee,Baron:2020xoi,Marzani:2017kqd}.
%
P-IVS approximates P-AS and preserves its continuity.
%
Finally, the Recursive Subtraction with a Fraction 1/2 (\PRSF{1/2}) algorithm we introduce in \Sec{rsf} furnishes, to our knowledge, the first example of a soft continuous tree-based jet grooming algorithm;
%
the soft discontinuities of other P-RS algorithms occur only in highly suppressed regions of phase space.

%--------------------------------
% Outline of the paper
%--------------------------------
The rest of the paper proceeds as follows:
%
in \Sec{PIRANHA}, we review the Energy Mover's Distance (EMD) of \Reff{Komiske:2019fks}, propose the \PIRANHA{} paradigm for continuous jet grooming, review the Apollonius Subtraction (P-AS) and Iterated Voronoi Subtraction (P-IVS) algorithms of \Reff{Komiske:2020qhg}, and introduce Recursive Subtraction (P-RS) with a Fraction (P-RSF) as a new, tree-based implementation of \PIRANHA{}.
%
In \Sec{traditionaldiscont}, we discuss the soft discontinuities present in hard-cutoff groomers and P-RSF, demonstrate the soft continuity of \PRSF{1/2}, and examine angular discontinuity in tree-based groomers.
%
In the remainder of the paper, we demonstrate the soft insensitivity enjoyed by \PIRANHA{} groomers over hard-cutoff groomers through several phenomenological studies.
%
In \Sec{soft_distortion}, we compare the responses of the hard-cutoff Soft Drop algorithm and \PRSF{1/2} to soft distortions from hadronization and to the removal of neutral particles, as a proxy for detector effects.
%
In \Sec{add_contam}, we compare the ability of Soft Drop and \PRSF{1/2} to effectively remove additive contamination from pileup and the underlying event.
%
We give concluding thoughts and avenues for future exploration in \Sec{Conclusions}, delegating discussions of \PIRANHA{} grooming in EMD mode, comparisons of other hard-cutoff and \PIRANHA{} groomers, and a first look at the perturbative structure of \PIRANHA{} grooming to \Appss{grooming_in_emd_mode}{feedingfrenzy}{calc}, respectively.
%
The \PIRANHA{} algorithms used in our analysis are available on GitHub \cite{piranhagithub}.

%%%%%%%%%%%%%%%%%%%%%%%%%%
% PIRANHA:
%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The PIRANHA{} Paradigm for Continuous Grooming}
\label{sec:PIRANHA}

This section introduces the \PIRANHA{} paradigm for jet grooming, which uses optimal transport to continuously remove low-energy pollution from jets.
%
We first review the Energy Mover's Distance (EMD) of Ref.~\cite{Komiske:2019fks}, which provides a framework to introduce the principles of \PIRANHA{} groomers and to enforce their continuity.
%
We then discuss three implementations of \PIRANHA{}:
%
Apollonius Subtraction (P-AS) and Iterated Voronoi Subtraction (P-IVS) -- both introduced in \Reff{Komiske:2020qhg} -- and the new tree-based Recursive Subtraction (P-RS) family of \PIRANHA{} algorithms.
%
We describe a variant of \PIRANHA{} -- grooming in EMD mode -- in \App{grooming_in_emd_mode}.

%=====================================
% EMD:
%=====================================
\subsection{Review of the Energy Mover's Distance}
\label{sec:emd}
The Energy Mover's Distance (EMD) of Ref.~\cite{Komiske:2019fks} provides a quantitative measure of the similarity between two jets.
%
We introduce the EMD here both to facilitate our definition of continuity in jet grooming and as a useful tool for quantifying the responses of jet groomers to low-energy pollution.
%
Readers familiar with the EMD should feel free to skip to \Sec{pira_intro}.

The EMD is a metric on the space of collider events.
%
Here and in the remainder of the paper, we borrow the terminology of \Reff{Komiske:2020qhg} and use the terms ``event'' and ``collider event'' to refer to the \textit{energy flow} of the event, or the angular distribution of energy of its outgoing radiation.\footnote{
For example, the energy flow of an event \(\mathcal E\) with \(M\) outgoing particles is the angular distribution
\begin{equation*}
    \label{eqn:energyflow}
    \mathcal E(\hat n)
    =
    \sum_{i=1}^ME_i\,\delta(\hat n - \hat n_i),
\end{equation*}
where $E_i > 0$ denotes the energy carried by particle \(i\), and $\hat n_i$ denotes its outgoing angular direction.
}
%
The EMD may be thought of as the amount of ``work'' required to rearrange one event into another.
%
For events that consist of a finite number of outgoing particles, the EMD is defined as the solution to the optimal transport problem
\begin{equation}
    \label{eqn:EMD_def_1}
    {\rm EMD}_{\beta, R}\left(\mathcal{E}, \mathcal{E}'\right)
    =
    \min_{f_{ij} > 0} \sum_{i = 1}^M \sum_{j = 1}^{M'} f_{ij} \left(\frac{\theta_{ij}}{R}\right)^\beta
    +
    \left| \sum_{i = 1}^M E_i -  \sum_{j = 1}^{M'} E'_j\right|,
\end{equation}
\begin{equation}
    \sum_{i = 1}^M f_{ij} \leq E_j',
    ~~
    \sum_{j = 1}^{M'} f_{ij} \leq E_i,
    ~~
    \sum_{i = 1}^M \sum_{j = 1}^{M'} f_{ij}
    = \min\left(\sum_{i = 1}^M E_i, \sum_{j = 1}^{M'} E'_j\right)
    \label{eqn:EMD_def_2}
    ,
\end{equation}
where \(i \in \{1, ..., M\}\) indicates a final-state particle of \(\mathcal{E}\) with energy \(E_i\), \(j \in \{1, ..., M'\}\) indicates a final-state particle of \(\mathcal{E}'\) with energy \(E'_j\), and \(\theta_{ij}\) denotes an angular distance between particles \(i\) and \(j\).
%
\(\beta > 0\) and \(R > 0\) are free parameters that control the behavior and relative importance of the first term on the right-hand side of \Eq{EMD_def_1}.\footnote{
Strictly speaking, the EMD defines a metric only if \(2 R\) is greater than the maximum possible angular distance \(\theta_{ij}\).
%
Furthermore, for \(\beta > 1\), one must raise the first term of \Eq{EMD_def_1} to the \(1/\beta\) power.
%
We use the EMD without a subscript, \(
{\rm EMD}\left(\mathcal{E}, \mathcal{E}'\right)
\) to denote the EMD with \(\beta = 1\) and with large enough \(R\) that it furnishes a metric.
}

In this work, we adopt the hadronic angular measure of \Reff{Komiske:2020qhg}, given by
\begin{equation}
    \theta_{ij} = \sqrt{-(n_i - n_j)^2}
    ,
    ~~~~~~
    ~~~~~~
    n_{i}^\mu = \frac{p_i^\mu}{E_{T\,i}} = (\cosh(y_i), \textbf{v}_{T\,i}, \sinh(y_i))^\mu
    .
    \label{eqn:hadronic_metric}
\end{equation}
Here, \(n_i^\mu\) is a vector describing the motion of particle \(i\), parameterized by its rapidity \(y_i\) and its velocity \(\textbf{v}_{T\, i}\) in the directions transverse to the beampipe.
%
This angular metric reproduces the rapidity-azimuth distance between particles \(i\) and \(j\) in the small rapidity-azimuth distance limit.\footnote{
Another popular choice for the form of the \(n_i^\mu\) is captured by
\begin{align}
    \label{eqn:ee_metric}
    n_{i,~{e^+e^-}}^\mu = \frac{p_i^\mu}{E_i} = (1, \textbf{v}_i)^\mu
    ,
\end{align}
which is a natural choice in the study of electron-positron collisions~\cite{Komiske:2020qhg}.
%
This choice reproduces the real-space opening angle between particles \(i\) and \(j\) in the small angle limit.
}

The EMD between a groomed jet in the presence and absence of low-energy pollution also provides an observable-independent tool for quantifying the sensitivity of jet grooming to pollution from soft distortions and additive contamination.
%
Since the EMD furnishes a metric on the space of events, two events are separated by zero EMD iff they have identical energy flow, and two jets separated by zero EMD will yield the same value for every infrared/collinear (IRC) safe observable (see \href{https://arxiv.org/pdf/2004.04159.pdf#page=11\&zoom=100,0,650}{Lemma 1} of \Reff{Komiske:2020qhg}).
%
Further, the EMD between two events also \textit{directly} bounds the differences in a large class of IRC-safe observable quantities (Lipschitz continuous observables) between the events~\cite{Komiske:2019fks}.
%
For example, the EMD between a parton-level jet and the same jet after hadronization (the parton-hadron EMD) directly bounds the difference between a large class of parton- and hadron-level IRC-safe observables.
%
The parton-hadron EMD is therefore a powerful, observable independent tool for characterizing the response of a jet to soft distortions from hadronization.
%
We use the EMD in our analysis of grooming sensitivity to several sources of low-energy pollution in \Secs{soft_distortion}{add_contam}.

%=====================================
% The PIRANHA Paradigm for Jet Grooming:
%=====================================
\subsection{PIRANHA{} and Continuity}
\label{sec:pira_intro}

\PIRANHA{} grooming may be intuitively imagined as the optimal transport of a group of hungry piranhas towards the low-energy pollution within a jet.
%
In this analogy, henceforth the piranha analogy, a \PIRANHA{} groomer populates the \(\eta\)-\(\phi\) plane with a school of piranhas, distributed according to a particular model of the polluting energy density.
%
When they are given an event to groom, the piranhas discuss which piranhas will get to eat which jet constituent;
%
in the language of optimal transport, we say that the piranhas decide on a set of transport plans.
%
The piranhas then implement the plan and feed on the constituents of the jet, or subtract the \(p_T\) of each jet constituent.

>>>>

\chapter{Removing the Rubble:\\Jet Substructure, Grooming, Pileup Mitigation, and Cleanup Fish}

% ==============================================
\section{The physical picture}
% ==============================================

% -----------------------------------
% Figure Comment:
% -----------------------------------
\begin{figure}[t!]
    % Figure graphics
    \centering
    \includegraphics[width=\textwidth]{figures/tempfig}

    % Caption
    \caption{
        Caption text
    }

    % Figure Label
    \label{fig:label}
\end{figure}
% -----------------------------------

% ==============================================
\section{Jet Substructure}
% ==============================================

% ==============================================
\section{Jets, Data, and Noise}
% ==============================================

% ---------------------------------------
\subsection{Jet Grooming and Pileup Mitigation}
% ---------------------------------------

% ==============================================
\section{Continuity as a bridge between grooming and pileup mitigation}
% ==============================================

% ==============================================
\section{Pileup and Infrared Annihilation (PIRANHA): A paradigm for continuous jet grooming}
% ==============================================

% ==============================================
\section{Alternatives to Grooming}
% ==============================================

% ==============================================
\section{Connecting back out}
% ==============================================
